---
title: 'Final Project: Academy Award Nominated Films, Violence, and the Bechdel Test'
output:
  html_document: default
  pdf_document: default
  word_document: default
date: "2026-05-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(readr)
library(ggplot2)

options(scipen = 999)
film_data <- read_csv('./data/final_project_data.csv')
  skip_empty_rows = TRUE
  stringsAsFactors = TRUE
```

### Esther Lim, Ethan Sanders, Kamilia Kwiatkowska

# Research Question

### Among the Oscar Academy Award Best Picture shortlisted films from 2020 to 2026, is there a significant relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets?

# Hypotheses

### Hypothesis 1 (H1)

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 2 (H2)

There is a relationship between a film being female-directed and non-female-directed and its odds of passing all three Bechdel Test criteria among Oscar Best Picture shortlisted films (2020-2026).

### Hypothesis 3 (H3)

There is a relationship between a film’s worldwide box office revenue and the IMDb Violence and Gore rating among Oscar Best Picture shortlisted films (2020-2026).

# Findings

## Hypothesis 1

There is a statistically detectable relationship between a film’s IMDb Violence and Gore rating and the number of Bechdel Test criteria it meets among Oscar Best Picture shortlisted films (2020-2026).

### Median and Mean (H1)

```{r}
violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

film_data %>%
  group_by(
    violence
  ) %>%
  summarize(
    n = n(),
    Mdn = median(bechdel),
    M = mean(bechdel),
    SD = sd(bechdel)
  )
```

### Scatter H1

```{r}
violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_point() +
  geom_jitter()
```

### Violin H1

```{r}
violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_violin()
```

### Boxplot H1

```{r}
violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = violence, y = bechdel)
  ) +
  geom_boxplot()
```

### Spearman Test H1

```{r}
film_data <- film_data %>%
  mutate(
    violence_coded = violence %>%
      as.numeric
  )

film_data %>%
  cor.test(
    ~ bechdel + violence_coded,
    data = .,
    method = 'spearman'
  )
```

### Findings H1

There is no real correlation between Bechdel Test criteria passed and violence scores for Academy Award Best Picture nominees, because p = 0.548. PLEASE CHECK.

# Hypothesis 2

H2 Of Best Picture nominees, winners of the Academy Award for Best Picture are more likely to pass Bechdel Test criteria.

### Median and Mean H2

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

film_data %>%
  group_by(
    result
  ) %>%
  summarize(
    n = n(),
    Mdn = median(bechdel),
    M = mean(bechdel),
    SD = sd(bechdel)
  )
```

### Scatter H2

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = result, y = bechdel)
  ) +
  geom_point() +
  geom_jitter()
```

### Violin H2

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = result, y = bechdel)
  ) +
  geom_violin()
```

### Boxplot H2

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = result, y = bechdel)
  ) +
  geom_boxplot()
```

### Spearman Test H2

```{r}
film_data <- film_data %>%
  mutate(
    result_coded = result %>%
      as.numeric
  )

film_data %>%
  cor.test(
    ~ bechdel + result_coded,
    data = .,
    method = 'spearman'
  )
```

### Findings H2

There is no real correlation between Bechdel Test criteria passed and winning the Academy Award Best Picture among Academy Award Best Picture nominees, because p = 0.139. PLEASE CHECK.

# Hypothesis 3

H3 Of Best Picture nominees, winners of the Academy Award for Best Picture are more likely to score a higher level of violence.

### Median and Mean H3

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

film_data %>%
  group_by(
    result
  ) %>%
  summarize(
    n = n(),
    Mdn = median(violence_coded),
    M = mean(violence_coded),
    SD = sd(violence_coded)
  )
```

### Scatter H3

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = result, y = violence)
  ) +
  geom_point() +
  geom_jitter()
```

### Violin H3

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = result, y = violence)
  ) +
  geom_violin()
```

### Boxplot H3

```{r}
result_levels <- c(
  'Winner',
  'Nominee'
)

film_data <- film_data %>%
  mutate(
    result = result %>%
      ordered(
        levels = result_levels
      )
  )

violence_levels <- c(
  'None',
  'Mild',
  'Moderate',
  'Severe'
)

film_data <- film_data %>%
  mutate(
    violence = violence %>%
      ordered(
        levels = violence_levels
      )
  )

film_data %>%
  ggplot(
    aes(x = result, y = violence)
  ) +
  geom_boxplot()
```

### Spearman Test H3

```{r}
film_data <- film_data %>%
  mutate(
    result_coded = result %>%
      as.numeric
  )

film_data <- film_data %>%
  mutate(
    violence_coded = violence %>%
      as.numeric
  )

film_data %>%
  cor.test(
    ~ violence_coded + result_coded,
    data = .,
    method = 'spearman'
  )
```

### Findings H3

There is no real correlation between violence score and winning the Academy Award Best Picture among Academy Award Best Picture nominees, because p = 0.533. PLEASE CHECK.

# Hypothesis 4

H4 The higher the Box Office revenue for an Academy Award for Best Picture nominated film, the more Bechdel Test criteria the film will pass.

### Median and Mean H4

```{r}
film_data %>%
  group_by(
    bechdel
  ) %>%
  summarize(
    n = n(),
    Mdn = median(box_office),
    M = mean(box_office),
    SD = sd(box_office)
  )
```

### Scatter H4

```{r}
film_data %>%
  ggplot(
    aes(x = box_office, y = bechdel)
  ) +
  geom_point() +
  geom_jitter()
```

### Violin H4

```{r}
film_data %>%
  ggplot(
    aes(x = box_office, y = bechdel)
  ) +
  geom_violin()
```

### Boxplot H4

```{r}
film_data %>%
  ggplot(
    aes(x = box_office, y = bechdel)
  ) +
  geom_boxplot()
```

### Spearman Test H4

```{r}
film_data %>%
  cor.test(
    ~ bechdel + box_office,
    data = .,
    method = 'spearman'
  )
```

### Findings H4

There is no real correlation between Bechdel Test criteria met and Box Office revenue for Academy Award Best Picture nominees, because p = 0.353. PLEASE CHECK.

# Hypothesis 5

H5 The higher the Box Office revenue for an Academy Award for Best Picture nominated film, the film will score a higher level of violence.

### Median and Mean H5

```{r}
film_data %>%
  group_by(
    violence
  ) %>%
  summarize(
    n = n(),
    Mdn = median(box_office),
    M = mean(box_office),
    SD = sd(box_office)
  )
```

### Scatter H5

```{r}
film_data %>%
  ggplot(
    aes(x = box_office, y = violence)
  ) +
  geom_point() +
  geom_jitter()
```

### Violin H5

```{r}
film_data %>%
  ggplot(
    aes(x = box_office, y = violence)
  ) +
  geom_violin()
```

### Boxplot H5

```{r}
film_data %>%
  ggplot(
    aes(x = box_office, y = violence)
  ) +
  geom_boxplot()
```

### Spearman Test H5

```{r}
film_data <- film_data %>%
  mutate(
    violence_coded = violence %>%
      as.numeric
  )

film_data %>%
  cor.test(
    ~ violence_coded + box_office,
    data = .,
    method = 'spearman'
  )
```

### Findings H5

There is no real correlation between violence score and Box Office revenue for Academy Award Best Picture nominees, because p = 0.354. PLEASE CHECK.

```{r}

bechdel_result_table <- table(film_data$result_coded, film_data$bechdel)

bechdel_result_odd <- fisher.test(bechdel_result_table)

bechdel_result_odd$estimate

```